---
title: "Assignment4"
format: html
editor: visual
---

Assignment 4, due Friday, x?

# **Instructions**

-   Total = 10 points + x bonus. Most questions are 0.5 point each.

-   Cite your sources by pasting links below the answers to each question. Acknowledge collaborators in the credits at the beginning.

**Goal:** This assignment is meant to make you get hands-on practice the R content at your own pace. Do take the time to review the slides, `.R` worksheet and re-watch lectures if you need, to recall topics mentioned in the course and the in-class activity.

**Format:** Please type out your answers within R chunks in this `.qmd` (quarto markdown) script, along with comments added below the code as plain text to briefly explain your answer / steps taken. \_If relevant, make short comment next to the line of code with `#…`. Please make sure to run and test the `R` chunks in RStudio before submission. If possible, try to `knit`/`Render` the document into the `.html` output format and submit this along with the `.qmd` file on Canvas. In short: **upload the `.qmd` script + `.html` output (*optional*) as a file attachments on canvas.**

**Bonus**: (***NEW***) Bonus 0.x series questions check for basic understanding/fundamentals/ topics mentioned previously as an opportunity for anyone who didn't understand these previously to catch up by reading on them and demonstrating learning. *If you already knew the answer well enough, you can skip these bonus questions and spend time more effectively on the harder bonus questions!* 😉

Other bonus questions make you go beyond the material covered in the class, and would require self exploration, i.e.) searching the Internet, parsing out correct answers, and understanding them enough to solve it. Hence *documenting your sources and proper explanation is very important for bonus questions*. **Any points beyond the maximum for in this assignment will carried over to your course grade**

## Scope

This assignment is meant to be a guide to exploring your data with tabular manipulations (`dplyr`) and visual inspection (`ggplot`). Read more about exploratory data analysis [here](https://r4ds.had.co.nz/exploratory-data-analysis.html). *The material in this homework was covered in these lectures -*

-   Tidyverse: class 6, 7

-   ggplot: class 8

## Resources for extra help

1.  Interactive tutorial to brush up `dplyr` functions such as `mutate()`, `filter()`, `arrange()` and others. I strongly recommend going through this exercise step-by-step for gaining confidence.
2.  Markdown formatting guide

## Credits for collaboration

*Please list your collaborators here with their key contribution summarized in a short sentence. For example -*

> **Annie**: Ideas for data-set and a few questions
>
> **Sam**: General discussion

# Questions

## Q0. Loading libraries and data

***Bonus 0.a.** (0.25 point each x 3)*

-   *Explain why we need to use* `library()` *command in one sentence with a citation;*

-   *where is* `install.packages()` *installing the gapminder package from?*

    -   *Which of these two commands do you require the internet to run?*

-   Complete the comments next to the code in the `R` chunk below

```{r libraries}
library(tidyverse) # add comment on package purpose..

install.packages('gapminder') # "
```

### Data for this work

1.  `midwest` : We will use the data set `midwest` containing demographic information of midwest counties from the 2000 US census. Reference: [ggplot2](https://ggplot2.tidyverse.org/reference/midwest.html)

2.  `gapminder` : This data contains the population and GDP per capita of multiple countries, and continents across years (1952 - 2007). Reference: [CRAN](https://cran.r-project.org/web/packages/gapminder/vignettes/gapminder.html)

### Loading data

1.  Load the data from the gapminder package (0.25 points)

-   Hint: answer is written below, but you need to explain the difference in the 3 `gapminder` words used here (***Bonus 0.b.**/+0.25 points)*

-   If that is too confusing, use your own method to load the data 🙂

```{r load_data}
gapminder <- gapminder::gapminder # what does this do?
```

2.  Load the midwest dataset from the ggplot2 package (0.25 points)

*Make your own code chunk and load the midwest dataset here*

------------------------------------------------------------------------

## Q1: Histograms

Create a 2 histograms, one showing `percbelowpoverty` and one showing `percollege` (% college educated) (*0.5 points)*

```{r histograms}

```

Display both graphs at the same time using the `patchwork` package (*0.25 points)*

*Here's hints for how to use [patchwork](https://patchwork.data-imaginist.com/articles/patchwork.html) to [assemble](https://patchwork.data-imaginist.com/articles/guides/assembly.html) plots*

```{r assemble_patchwork}

```

*Describe the patterns in the histograms in one short sentence each.* (0.25 points)

(*Rubric*) '\[0.25 points for each histogram\] ; '\[-0.25 if they don't explain the pattern they see\]

## Q2: Scatter plot

You now wish to explore the relationship between poverty and college degrees. To do this, we create a scatterplot of `percbelowpoverty` vs. `percollege`. Color the datapoints by the `inmetro` column (*hint: use as.factorO, and write what a 0 means and what a 1 means*) (0.5 points)

```{r scatterplot}

```

Change the shape of data points based on the state. (0.25 points)

```{r scatter_size}

```

Create another scatterplot, this time adding adding a geom_smooth line. (0.25 points)

```{r smooth_fit}

```

Describe the pattern you see with the data. What story is the data telling? (*0.25 points*)

### Q2.b (bonus): Custmize legend 

Rename the legend for `inmetro` to be more intuitive than 0 & 1 (*0.25 points*)

```{r legend_gymnastics}

```

(*Rubric*)

'\[-0.25 if they color points incorrectly\] ; '\[-0.5 if they only do 1 histogram\] ; '\[-0.5 if incorrect columns used\] ; '\[-0.25 if they don't change inmetro to a factor\] ; '\[-0.5 if they dont explain the pattern they see\]

## Q3: barplots

Find the average popdensity across all counties. Create a new column that tells if a county is above or below average population density using `mutate()`. *(name it something like above_density).* (*0.25 point*)

```{r data_wrangling}

```

Create a barplot that shows how many counties are above and below the average. *Hint: you should have 2 boxes (0.25 point)*

```{r barplot}

```

Color the counts in the bars by state. *(0.25 point)*

```{r colourbarplot}

```

### Q3.b (bonus): mean for each state 

Find the mean for EACH STATE and repeat the above question, with rhe above_mean_density specific to the state the county is found in. (*2 points)*

*Hint:* need to use `group_by()`

(*Rubric*) : '\[-0.25 if calculate the mean incorrectly\] ; '\[-0.25 if they do not color by state\] ; '\[-0.5 if barplot is incorrect\]

------------------------------------------------------------------------

## Q4: Inspect the gapminder data

When you get data, it is a good idea to know the column names, approximate size of data (rows and columns).

What is one of the R command that let's you do this?

```{r know_thy_data}

```

## Q5: Visually inspect the gapminder data

Play with the data and make 2 timeseries plots to get a feel for your your favorite country's data over time *(x axis is time variable, y axis anything else)*

```{r visual_eda1}


```

------------------------------------------------------------------------

# Attributions

[Gapminder](https://cran.r-project.org/web/packages/gapminder/index.html) dataset.
